Towards a Machine Learning approach for Aggregate Computing

### 2021-11-28 paper-2021-woa-ac-ml.tex
Oxford comma needed
Explain hopCount in comments + minors in code style

Reward function instead, might be depending on the learning technique, time and problem. 
 Indeed, if we perform learning offline, it could be defined from the overall \textit{simulated} computational field:
$$
\textit{R}_{global} := \psi({\theta})
$$
If we use an online mode, it needs to be defined as a function to the local node context:
$$
\textit{R}_{local} := \psi(\textit{Context})
$$
Finally, actions influence the current aggregate program output, changing the program evaluation as:
$$
f_{rl} : \textit{Context} \rightarrow \lambda(\textit{Output}, \textit{Action})
$$
As remarked before, the policy $\pi$ need to be distributed, so each agent can choose the action that will be performed:
$$
\textit{Action} \leftarrow \pi(K)
$$

@@

A naive implementation of this kind of problem can be described in AC as:
\begin{minted}[fontsize=\scriptsize]{scala}
def hopCount(): Int = {
  rep(Double.PositiveInfinity) { 
    hopCounts => mux(source) 
    { 0 } 
    { minHoodPlus(nbr{hopCounts}) + 1 }
  }
}
\end{minted}
However, this solution suffers from the slow-rising problem~\cite{DBLP:conf/saso/AudritoCDV17}. 
 It consists of a slow convergence time when an equilibrium situation is broken.
So, our first intuition is to accelerate convergence through learning.
 In particular, the previous solution can be rewritten as:

 \begin{minted}[fontsize=\scriptsize]{scala}
def hopCountRl(k: Int, w: Int): Int = 
{
  rep(Double.PositiveInfinity) { 
    hopCounts => 
      val currentState = 
        state(hopCounts, k, w)
      val action = localPolicy(state)
      hopCout() + action
  }
}
\end{minted}

>>>>

Reward function instead, might be depending on the learning technique, time, and problem. 
 Indeed, if we perform learning offline, it could be defined from the overall \textit{simulated} computational field:
$$
\textit{R}_{global} := \psi({\theta})
$$
If we use an online mode, it needs to be defined as a function to the local node context:
$$
\textit{R}_{local} := \psi(\textit{Context})
$$
Finally, actions influence the current aggregate program output, changing the program evaluation as:
$$
f_{rl} : \textit{Context} \rightarrow \lambda(\textit{Output}, \textit{Action})
$$
As remarked before, the policy $\pi$ need to be distributed, so each agent can choose the action that will be performed:
$$
\textit{Action} \leftarrow \pi(K)
$$

@@

A naive implementation of this kind of problem can be described in AC as:
\begin{minted}[fontsize=\scriptsize,mathescape]{scala}
def hopCount(): Int = 
  // update state (initially, $\infty$; now, hopCounts) by ...
  rep(Double.PositiveInfinity) { hopCounts => 
    // source returns 0,
    // non-sources minimise over neighbourdhood and add 1
    mux(source) { 0 }{ minHoodPlus(nbr{hopCounts}) + 1 }
  }
\end{minted}
However, this solution suffers from the slow-rising problem~\cite{DBLP:conf/saso/AudritoCDV17}. 
 It consists of a slow convergence time when an equilibrium situation is broken.
So, our first intuition is to accelerate convergence through learning.
 In particular, the previous solution can be rewritten as:

 \begin{minted}[fontsize=\scriptsize]{scala}
def hopCountRl(k: Int, w: Int): Int = 
  rep(Double.PositiveInfinity) { hopCounts => 
      val currentState = 
        state(hopCounts, k, w)
      val action = localPolicy(state)
      hopCout() + action
  }
\end{minted}
